NDNS multilevel LCA CW2CB2 or CW2CB3

### 2018-08-15 paper/Chapters/Chapter3.tex
discussion updated

Class 3 days \textbf{(Figure \ref{fig:level1}-C)} were called "regular meals day" due to the following reasons: 1) participants' dietary recordings showed that in these days there was almost 0 possibility of not eating any food at lunch (0.019 between 12 noon and 2 pm) and dinner (0.034 between 5 pm and 8 pm); 2) the probabilities of not eating during morning snack time (9 am to 12 am) and afternoon snack time (2 pm to 5 pm) were also the highest among the three types of days (0.401 and 0.659). 3) during these days, participants may have some high carbohydrate contained food between 8 pm and 10 pm (probability = 0.308), but the probability of not eating any food during 10 pm to 6 am next morning was 0.751, the highest among the three types of days. \vspace{-0.5cm}

@@

Among the three types of carbohydrate eaters, the mean of total energy intake over the 4 days of dietary survey was the highest (7985.8 kJ, 95\%CI: 7283.3, 8146.3) in the low carbohydrate eaters group, and the lowest (7341.8 kJ, 95\%CI: 7172.5, 7511.2) in the moderate eaters group \textbf{(Table \ref{tab:tab-nutri-indi})}. Sources of energy for each type of carbohydrate eaters by the 7 time slots were also different. Low carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-A)} never had carbohydrate contributed more than 50\% of their total energy throughout the day. Energy from fat were the highest for low carbohydrate eaters most of the time during the day (except for time between 10 pm to 6 am next morning). Most impressively, energy from alcohol were always the highest in low carbohydrate eaters, percentages for energy from alcohol for the 7 times slots were 1.4\% (6-9 am), 1.8\% (9-12 am), 3.1\% (12-2 pm), 7.2\% (2-5 pm), 8.4\% (5-8 pm), 20.0\% (8-10 pm), and 31.2\% (10 - 6 am), respectively. Contribution from different energy sources are quite similar for moderate and high carbohydrate eaters, but their absolute amount of energy consumption at each time slot were largely different. Moderate carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-B)} were characterised as consuming the lowest energy (296.9 kJ) before 9 am, but having higher energy consumption (967.9 kJ) between 9 am and 12 noon time compared with low and high carbohydrate eaters. Moderate carbohydrate eaters may turn to have later breakfast, later lunch, and probably later dinner as well. They had the highest total energy consumption (599.7 kJ) at night (10 pm - 6 am) across three types of eaters. High carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-C)} consumed the highest total energy (929.0 kJ) during 6 am to 9 am in the morning and the lowest total energy between 10 pm to 6 am (205.5 kJ). Specifically, carbohydrate contribution to total energy intake were 65.3\% (6-9 am), 56.2\% (9-12 am), 46.2\% (12-2 pm), 50.3\% (2-5 pm), 42.5\% (5-8 pm), 44.1\% (8-10 pm), and 49.9\% (10-6am). We also noticed that high carbohydrate eaters consumed their energy mainly from three time slots: 6-9 am, 12-2 pm, and 5-8 pm.

As expected, in total, the mean of carbohydrate intake was 203.8 g, 218.3 g, and 233.4 g for low, moderate, and high carbohydrate eaters, respectively \textbf{(Table \ref{tab:tab-nutri-indi})}. Energy contribution from carbohydrate was close to 50\% in the high carbohydrate eaters, but was only 40.6\% in the low carbohydrate eaters. In terms of the subtypes (components) of the carbohydrate consumed at each time slot, high carbohydrate eaters consumed as much as more than 2 times (compared to low carbohydrate eaters) and nearly 4 times (against moderate carbohydrate eaterss) the amount of sugar (37.9g 95\% CI: 36.8, 39.2) and non-milk extrinsic sugar (i.e. free sugar) (11.1g 95\%CI: 10.7, 11.6) between 6-9 am. Moderate carbohydrate eaters had their carbohydrate intake more spread out. They consumed more sugar and starch during 9-12 am, 2-5 pm, 8-10 pm, and 10-6 am. Low carbohydrate eaters turned to have similar temporal pattern of consuming carbohydrates but the absolute amount of fibre, sugar, free sugar, and starch were usually lower than that in the high carbohydrate eaters except for time slots of 2-5 pm, and 10-6 am. Strong evidence (\textit{p} < 0.001) suggested that the mean of total fibre consumption for low,

@@

		\hspace{1em}\textbf{9 am – 12 am} & 25.1 (23.9, 26.3) & 33.0 (31.4, 34.6) & 26.2 (25.1, 27.2) & \\

>>>>

Class 3 days \textbf{(Figure \ref{fig:level1}-C)} were called "regular meals day" due to the following reasons: 1) participants' dietary recordings showed that in these days there was almost 0 possibility of not eating any food at lunch (0.019 between 12 noon and 2 pm) and dinner (0.034 between 5 pm and 8 pm); 2) the probabilities of not eating during morning snack time (9 am to 12 noon) and afternoon snack time (2 pm to 5 pm) were also the highest among the three types of days (0.401 and 0.659). 3) during these days, participants may have some high carbohydrate contained food between 8 pm and 10 pm (probability = 0.308), but the probability of not eating any food during 10 pm to 6 am next morning was 0.751, the highest among the three types of days. \vspace{-0.5cm}

@@

Among the three types of carbohydrate eaters, the mean of total energy intake over the 4 days of dietary survey was the highest (7985.8 kJ, 95\%CI: 7283.3, 8146.3) in the low carbohydrate eaters group, and the lowest (7341.8 kJ, 95\%CI: 7172.5, 7511.2) in the moderate eaters group \textbf{(Table \ref{tab:tab-nutri-indi})}. Sources of energy for each type of carbohydrate eaters by the 7 time slots were also different. Low carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-A)} never had carbohydrate contributed more than 50\% of their total energy throughout the day. Energy from fat were the highest for low carbohydrate eaters most of the time during the day (except for time between 10 pm to 6 am next morning). Most impressively, energy from alcohol were always the highest in low carbohydrate eaters, percentages for energy from alcohol for the 7 times slots were 1.4\% (6-9 am), 1.8\% (9-12 noon), 3.1\% (12-2 pm), 7.2\% (2-5 pm), 8.4\% (5-8 pm), 20.0\% (8-10 pm), and 31.2\% (10 - 6 am), respectively. Contribution from different energy sources are quite similar for moderate and high carbohydrate eaters, but their absolute amount of energy consumption at each time slot were largely different. Moderate carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-B)} were characterised as consuming the lowest energy (296.9 kJ) before 9 am, but having higher energy consumption (967.9 kJ) between 9 am and 12 noon time compared with low and high carbohydrate eaters. Moderate carbohydrate eaters may turn to have later breakfast, later lunch, and probably later dinner as well. They had the highest total energy consumption (599.7 kJ) at night (10 pm - 6 am) across three types of eaters. High carbohydrate eaters \textbf{(Figure \ref{fig:energysourcesCB}-C)} consumed the highest total energy (929.0 kJ) during 6 am to 9 am in the morning and the lowest total energy between 10 pm to 6 am (205.5 kJ). Specifically, carbohydrate contribution to total energy intake were 65.3\% (6-9 am), 56.2\% (9-12 noon), 46.2\% (12-2 pm), 50.3\% (2-5 pm), 42.5\% (5-8 pm), 44.1\% (8-10 pm), and 49.9\% (10-6am). We also noticed that high carbohydrate eaters consumed their energy mainly from three time slots: 6-9 am, 12-2 pm, and 5-8 pm.

As expected, in total, the mean of carbohydrate intake was 203.8 g, 218.3 g, and 233.4 g for low, moderate, and high carbohydrate eaters, respectively \textbf{(Table \ref{tab:tab-nutri-indi})}. Energy contribution from carbohydrate was close to 50\% in the high carbohydrate eaters, but was only 40.6\% in the low carbohydrate eaters. In terms of the subtypes (components) of the carbohydrate consumed at each time slot, high carbohydrate eaters consumed as much as more than 2 times (compared to low carbohydrate eaters) and nearly 4 times (against moderate carbohydrate eaterss) the amount of sugar (37.9g 95\% CI: 36.8, 39.2) and non-milk extrinsic sugar (i.e. free sugar) (11.1g 95\%CI: 10.7, 11.6) between 6-9 am. Moderate carbohydrate eaters had their carbohydrate intake more spread out. They consumed more sugar and starch during 9-12 noon, 2-5 pm, 8-10 pm, and 10-6 am. Low carbohydrate eaters turned to have similar temporal pattern of consuming carbohydrates but the absolute amount of fibre, sugar, free sugar, and starch were usually lower than that in the high carbohydrate eaters except for time slots of 2-5 pm, and 10-6 am. Strong evidence (\textit{p} < 0.001) suggested that the mean of total fibre consumption for low,

@@

		\hspace{1em}\textbf{9 am – 12 noon} & 25.1 (23.9, 26.3) & 33.0 (31.4, 34.6) & 26.2 (25.1, 27.2) & \\

### 2018-08-15 paper/Chapters/Chapter4.tex
discussion updated

Among the dietary diaries collected, there were three distinct latent classes specifically for carbohydrate intake: 1) high probabilities of having high carbohydrate contained food across the hours of day (high percentage carbohydrate day); 2) low carbohydrate food dominant through out the hours of day (low percentage carbohydrate day); and 3) always having lunch and dinner day (regular meals day). And from these day level classification and their features, one might anticipate that individuals who managed to follow the regular meals day might be eating a relatively healthier diet because it seemed to be a regular temporal eating habit; at this time point, we also believe that those who follow either high or low percentages would probably consume total energy.

However, when the MLCA extended the model to individual level, three types of persons are further defined depending on their 4-day-diary: 1) low carbohydrate eaters, who mostly followed "regular meals day"; 2) moderate carbohydrate eaters, who had similar probabilities of following either "regular meals day" or "high percentage carbohydrate day"; 3) high carbohydrate eaters, who followed "high percentage carbohydrate day" for half of their survey. For the first time, as far as we know, the day-to-day food intake pattern variation within individuals was successfully captured. Results from the MLCA models showed that from the perspective of carbohydrate consumption, people were indeed switching between different types of carbohydrate eating types even within a short term period of survey. The MLCA models allowed the probability of a certain type of carbohydrate eating day to vary across individuals. This properly accounted for the fact that for some people, their probability of following a type of food eating pattern during the survey was higher/lower than that in the others. Apparently, assuming that a person is always following or characterised by one type of food eating pattern is not appropriate.

Surprisingly, low carbohydrate eaters turned out to consume the highest amount of total energy,

>>>>

Among the dietary diaries collected, there were three distinct latent classes specifically for carbohydrate intake: 1) high probabilities of having high carbohydrate contained food across the hours of day (high percentage carbohydrate day); 2) low carbohydrate food dominant through out the hours of day (low percentage carbohydrate day); and 3) always having lunch and dinner day (regular meals day). And from these day level classifications and their features, one might anticipate that individuals who managed to follow the regular meals day might be eating a relatively healthier diet because it seemed to be a regular temporal eating habit; at this time point, we also believed that those who followed either high or low carbohydrate percentage days would probably consume higher total energy than those who followed most regular meals days.

However, when the MLCA extended the model to individual level, three types of persons were further defined depending on their 4-day-diary: 1) low carbohydrate eaters, who mostly followed "regular meals day"; 2) moderate carbohydrate eaters, who had similar probabilities of following either "regular meals day" or "high percentage carbohydrate day"; 3) high carbohydrate eaters, who followed "high percentage carbohydrate day" for half of their survey. For the first time, as far as we know, the day-to-day food intake pattern variation within individuals was successfully captured by MLCA models. Results from the MLCA models showed that from the perspective of carbohydrate consumption, people were indeed changing their diet from day to day even within a short term period of survey. The MLCA models allowed the probability of following a certain type of carbohydrate eating day to vary across individuals. This properly accounted for the fact that for some people, their probability of following a type of food eating pattern during the survey could be higher/lower than that in the others. This finding also suggested that assuming a person will following a certain type of food intake pattern is not appropriate.

Surprisingly, low carbohydrate eaters whose dietary recordings suggested that they were mostly following a regular temporal meals pattern turned out to consume the highest amount of total energy among the three type of carbohydrate eaters. Detailed profiling of energy composition according to the time slots revealed that low carbohydrate eaters actually had higher proportion of energy contributed by both alcohol and fat. High percentage of fat consumption was shown in all 7 time slots, energy coming from alcohol exceeded more than one fifth of the total energy after 8 pm. These findings explained why they were actually consuming the highest energy among the three types of carbohydrate eaters. However, we also found that participants consuming low carbohydrate had higher prevalence of diabetes, hypertension, and obesity. These health issues might possibly lead them (or advices were given from their physicians) to replace carbohydrates in their diet to other energy sources such as fat, protein, or even alcohol. Therefore, there is a possibility that they chose to follow low carbohydrate diets out of healthy purpose, but many of them were replacing carbohydrates with higher energy condensed food or even alcohol at night which might indeed be a public health concern.

Further more, looking into the details of the timing and composition of the energy intake among the moderate carbohydrate eaters. We realised that although moderate carbohydrate eaters did not consume as much alcohol as low carbohydrate eaters at night. But they were
